---
title: "Analysising CPUE"
author: "Georgina Shaw"
date: "27/10/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```



## Comparing the median CPUE to the mean CPUE, incuding a time series

Firstly load all the packages needed
```{r, eval=TRUE}
library(readr)
library(dplyr)
library(lubridate)
library(magrittr)
library(tidyverse)
library(ggplot2)
```

Next read in the data 
```{r, eval=TRUE}
landings_data <- read_csv("sample_landings_data_raw.txt")
```

Renaming the columns and turning the date column into a date format that R recognises
```{r, eval=TRUE}
landings_data <- landings_data %>%
  rename(Year = yy,
         Date = dat,
         Trip_ID = trip,
         Effort_Hours = effort,
         Gear = gr,
         Species = sp,
         Length_cm = l_cm,
         Weight_g = w_cm) %>%
   mutate(Date = mdy(Date)) 
```  



I cleaned the data and removed any values with NA so it would be easier to calculate mean and medians 
```{r, eval=TRUE}
landings_data[!complete.cases(landings_data),]
landings_data <- na.omit(landings_data)
```

Computing the CPUE 
```{r, message=FALSE, eval=TRUE}
cpue_data <- landings_data %>% 
    mutate(Weight_kg = Weight_g / 1000) %>%
    group_by(Year,Trip_ID) %>% 
    summarize(Trip_CPUE = sum(Weight_kg) /           mean(Effort_Hours)) %>% 
    group_by(Year) %>% 
    summarize(Median_CPUE_kg_hour =                  median(Trip_CPUE), Mean_CPUE_kg_hour =           mean(Trip_CPUE), Trip_CPUE = Trip_CPUE) 
```
```{r}
cpue_data
```

Computing a graph comparing median CPUE to the mean CPUE with indication of the year they are from
```{r, eval=TRUE, message=TRUE}
  ggplot(data = cpue_data, mapping =  aes(x=Median_CPUE_kg_hour,y=Mean_CPUE_kg_hour)) +
  geom_point(mapping = aes(x = Median_CPUE_kg_hour, y = Mean_CPUE_kg_hour, color = Year)) +
  geom_line() + 
  ylab("Mean CPUE [kg/hour]") + xlab("MEdian CPUE [kg/houe]") +
  ggtitle("Comparing Median CPUE against Mean CPUE") 
```

Boxplot of CPUE
```{r}
boxplot(cpue_data$Trip_CPUE)
```