---
title: "Applying Gaussian Mixture Model"
author: "Georgina Shaw"
date: "07/03/2022"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("data/stomach_dataset.Rdata")
library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
```

# Herrin 


Only keeping the information we need 
```{r}
df <- stom_df %>%
  transmute(Species = pred_species,
            wprey = prey_weight_g,
            wpredator = pred_weight_g,
            Nprey = prey_count / n_stomachs,
            l = log(wpredator / wprey))
```
Choosing the type of fish to use, this time looking at cod 
```{r}
stomach <- df %>%
  filter(Species == "Clupea harengus",
         wprey > 0)

stomach %>% 
  group_by(Species) %>% 
  summarise(wprey_min = min(wprey),
            wprey_max = max(wprey),
            lmin = min(l),
            lmax = max(l))

stomach %>% 
  group_by(Species) %>% 
  filter(wprey == max(wprey))
```
Creating bins for the data 
```{r}
no_bins <- 30  # Number of bins
binsize <- (max(stomach$l) - min(stomach$l)) / (no_bins - 1)
breaks <- seq(min(stomach$l) - binsize/2,
              by = binsize, length.out = no_bins + 1)
```
Splitting the data into the bins that have been made
```{r}
binned_stomach <- stomach %>% 
  # bin data
  mutate(cut = cut(l, breaks = breaks, right = FALSE,
                   labels = FALSE)) %>% 
  group_by(Species, cut) %>% 
  summarise(Numbers = sum(Nprey), 
            Biomass = sum(Nprey * wprey)) %>% 
  # normalise
  mutate(Numbers = Numbers / sum(Numbers) / binsize,
         Biomass = Biomass / sum(Biomass) / binsize)  %>%
  # column for predator/prey size ratio
  mutate(l = map_dbl(cut, function(idx) breaks[idx] + binsize/2))
binned_stomach
```
We convert this into the long table format preferred by ggplot2.
```{r}
binned_stomach <- binned_stomach %>%
  gather(key = "Type", value = "Density", Numbers, Biomass)
```
       
## Histograms 
Plot the histogram that represents estimates of the normalised number density and the normalised biomass density 
```{r}
binned_stomach %>% 
  ggplot(aes(l, Density, fill = Type)) +
  geom_col(position = "dodge") +
  facet_grid(rows = vars(Species), scales = "free_y") +
  xlab("Log of Predator/Prey mass ratio") +
  expand_limits(x = c(0, 30))
```
          
## Kernal Density Estimation 
```{r}
adjust <- 1/2  # decrease bandwidth for kernel estimate
stomach <- stomach %>% 
  group_by(Species) %>% 
  mutate(weight_numbers = Nprey / sum(Nprey),
         weight_biomass = Nprey * wprey / sum(Nprey * wprey))
ggplot(stomach) +
  geom_density(aes(l, weight = weight_numbers,
                   fill = "Numbers"),
               adjust = adjust) +
  geom_density(aes(l, weight = weight_biomass,
                   fill = "Biomass"),
               adjust = adjust) +
  facet_grid(rows = vars(Species), scales = "free_y") +
  xlab("Log of predator/prey mass ratio") +
  expand_limits(x = c(0, 29))
```
         
## Gaussian Distribution fit
Plotting the normal distribution for numbers 
```{r}
weighted.sd <- function(x, w) {
  sqrt(sum(w * (x - weighted.mean(x, w))^2))
}
fit <- stomach %>% 
  summarise(mean = weighted.mean(l, weight_numbers),
            sd = weighted.sd(l, weight_numbers))
stomach %>% 
  ggplot() +
  geom_density(aes(l, weight = weight_numbers),
               fill = "#00BFC4", adjust = adjust) +
  xlab("Log of Predator/Prey mass ratio") +
  stat_function(fun = dnorm, 
                args = list(mean = fit$mean, 
                            sd = fit$sd), 
                colour = "blue") +
  expand_limits(x = c(6, 30))
```
         
Now adding biomass
```{r}
stomach %>% 
  ggplot() +
  geom_density(aes(l, weight = weight_numbers,
                   fill = "Numbers"),
               adjust = adjust) +
  geom_density(aes(l, weight = weight_biomass,
                   fill = "Biomass"),
               adjust = adjust) +
  xlab("Log of Predator/Prey mass ratio") +
  stat_function(fun = dnorm, 
                args = list(mean = fit$mean, 
                            sd = fit$sd), 
                colour = "blue") +
  stat_function(fun = dnorm, 
                args = list(mean = fit$mean - fit$sd^2, 
                            sd = fit$sd), 
                colour = "red") +
  expand_limits(x = c(0, 30))
```

```{r}
weighted.sd <- function(x, w) {
  sqrt(sum(w * (x - weighted.mean(x, w))^2))
}
fit <- stomach %>% 
  summarise(mean = weighted.mean(l, weight_numbers),
            sd = weighted.sd(l, weight_numbers))
stomach %>% 
  ggplot() +
  geom_density(aes(l, weight = weight_numbers,
                   fill = "Numbers"),
               adjust = adjust) +
  geom_density(aes(l, weight = weight_biomass,
                   fill = "Biomass"),
               adjust = adjust) +
  xlab("Log of Predator/Prey mass ratio") +
  stat_function(fun = dnorm, 
                args = list(mean = fit$mean, 
                            sd = fit$sd), 
                colour = "blue") +
  stat_function(fun = dnorm, 
                args = list(mean = fit$mean - fit$sd^2, 
                            sd = fit$sd), 
                colour = "red") +
  expand_limits(x = c(0, 30))
```

## Gaussaian Mixture Model 
The Gaussian Mixture model is the plotting of multiple Gaussian distributions on one plot. The estimates of the parameters are found using expectation maximization (EM), which consists of two steps, the expectation step (E step) and the maximization step (M step). 

For this analysis only the the univariant case needs to be considered, with parameters $\mu_{k}$ and $\sigma_{k}$ for each k-th component, with $k=2$. The mixture component weightings are defined as $\phi_{k}$ with $\sum_{i=1}^K \phi_{i} = 1$, so the probabilities add to one. 

The E step involves calculating the expectation of of the the assignment of each class for each data point given the model parameters $\phi_{k}$, $\mu_{k}$ and $\sigma_{k}$. The M step involves maximization the expectations calculated in the E step, which updates the values $\phi_{k}$, $\mu_{k}$ and $\sigma_{k}$. Eventually this should converge, giving a maximum likelihood estimate of the parameters. 


If the Gaussian Mixture Model fitted the numbers density then is would be distributed by two Gaussians models such as;
$$
n_1(p) \propto \exp\left(-\frac{(p-\mu_{1})^2}
  {2\sigma_{1}^2}\right),
$$
and 
$$
n_2(p) \propto \exp\left(-\frac{(p-\mu_{2})^2}
  {2\sigma_{2}^2}\right).
$$
Then the biomass density would be given by a two normal distributions but with the means shifted by $\sigma^2$, which would give the following; 
$$b_1(p) \propto \exp\left(-\frac{(p-(\mu_{1}-\sigma_{1}^2)^2}
  {2\sigma_{1}^2}\right)$$
  
$$b_2(p) \propto \exp\left(-\frac{(p-(\mu_{2}-\sigma_{2}^2)^2}
  {2\sigma_{2}^2}\right)$$
  
Firstly, find the numbers and biomass PPMR for each observation using the definition of each;
$$
r^{num}_i = \frac{1}{n}\sum_{j=1}^n \frac{M_i}{m_j}
$$
$$
r^{bio}_{i} = \frac{M_i}{\frac{1}{n}\sum_{j = 1}^{n}m_{j}}
$$


Applying Gaussian Mixture Models in different ways 
First way by calculating each individual PPMR by numbers and biomass and then applying the EM algorithm to find the estimated parameters and then plotting 
```{r}
ppmr <- stomach %>%
  mutate(numbers = log(wpredator/(wprey * sum(Nprey))),
         biomass = log(wpredator*sum(Nprey)/wprey)) 
library(mixtools)
my_mix1 <- normalmixEM(ppmr$numbers, k = 2) #applied the EM algorithm finding the estimates of the parameters 
my_mix2 <- normalmixEM(ppmr$biomass, k = 2)
 
#plot of numbers 
ggplot(ppmr, aes(x = numbers)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Clupea harengus") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix1[["mu"]], #mean
    sd = my_mix1[["sigma"]], #standard deviation
    lambda = my_mix1[["lambda"]], #amplitude
    n = length(ppmr$numbers), #sample size
    binwidth = binsize #binwidth used for histogram
  )
#plot of biomass
ggplot(ppmr, aes(x = biomass), main ="Clupea harengus") +
  geom_histogram(binwidth = binsize) +
  ggtitle("Clupea harengus") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix2[["mu"]], #mean
    sd = my_mix2[["sigma"]], #standard deviation
    lambda = my_mix2[["lambda"]], #amplitude
    n = length(ppmr$biomass), #sample size
    binwidth = binsize #binwidth used for histogram
  )
```

Looking at Gaussian distribution for different species 
```{r}
stomach_all <- stom_df %>%
  select(Species = pred_species,
            wprey = prey_weight_g,
            wpredator = pred_weight_g) %>%
  group_by(Species) %>%
  filter(n() > 1000, wprey > 0) %>%
  mutate(Nprey = 1,
         l = log(wpredator / wprey),
         weight_numbers = Nprey / sum(Nprey),
         weight_biomass = Nprey * wprey / sum(Nprey * wprey))
unique(stomach_all$Species)        

no_bins <- 30  # Number of bins
binsize <- (max(stomach_all$l) - min(stomach_all$l)) / (no_bins - 1)
breaks <- seq(min(stomach_all$l) - binsize/2,
              by = binsize, length.out = no_bins + 1)

binned_stomach <- stomach_all %>% 
  # bin data
  mutate(cut = cut(l, breaks = breaks, right = FALSE,
                   labels = FALSE)) %>% 
  group_by(Species, cut) %>% 
  summarise(Numbers = sum(Nprey), 
            Biomass = sum(Nprey * wprey)) %>% 
  # normalise
  mutate(Numbers = Numbers / sum(Numbers) / binsize,
         Biomass = Biomass / sum(Biomass) / binsize)  %>%
  # column for predator/prey size ratio
  mutate(l = map_dbl(cut, function(idx) breaks[idx] + binsize/2)) %>% 
  gather(key = "Type", value = "Density", Numbers, Biomass)   
```
Plotting the graphs 
```{r}
grid <- seq(0, max(stomach_all$l), length = 100)
normaldens <- plyr::ddply(stomach_all, "Species", function(df) {
  data.frame( 
    l = grid,
    density = dnorm(grid, mean(df$l), sd(df$l))
  )
})

ggplot(stomach_all) +
  geom_density(aes(l, weight = weight_numbers), fill = "#00BFC4") +
  facet_wrap(~Species, scales = "free_y", ncol = 4) +
  xlab("Log of predator/prey mass ratio")  +
  geom_line(aes(l, density), data = normaldens,
            colour = "blue")

grid <- seq(0, max(stomach_all$l), length = 100)
shifted_normaldens <- plyr::ddply(stomach_all, "Species", function(df) {
  data.frame( 
    l = grid,
    density = dnorm(grid, mean(df$l) - sd(df$l)^2, sd(df$l))
  )
})

ggplot(stomach_all) +
  geom_density(aes(l, weight = weight_biomass), fill = "#F8766D") +
  facet_wrap(~Species, scales = "free_y", ncol = 4) +
  xlab("Log of predator/prey mass ratio") +
  ylab("Biomass density") +
  geom_line(aes(l, density), data = shifted_normaldens,
            colour = "red") 
```

    
Scomber scombrus
```{r, echo = FALSE}
ppmr2 <- stomach_all %>%
  mutate(numbers = log(wpredator/(wprey * sum(Nprey))),
         biomass = log(wpredator*sum(Nprey)/wprey)) %>%
  filter(Species == "Scomber scombrus")


my_mix1 <- normalmixEM(ppmr2$numbers, k = 2)
my_mix2 <- normalmixEM(ppmr2$biomass, k = 2)

#plot of numbers 
ggplot(ppmr2, aes(x = numbers)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Scomber scombrus") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix1[["mu"]], #mean
    sd = my_mix1[["sigma"]], #standard deviation
    lambda = my_mix1[["lambda"]], #amplitude
    n = length(ppmr2$numbers), #sample size
    binwidth = binsize #binwidth used for histogram
  )
#plot of biomass
ggplot(ppmr2, aes(x = biomass)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Scomber scombrus") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix2[["mu"]], #mean
    sd = my_mix2[["sigma"]], #standard deviation
    lambda = my_mix2[["lambda"]], #amplitude
    n = length(ppmr2$biomass), #sample size
    binwidth = binsize #binwidth used for histogram
  )
```
 
Limanda limanda 
```{r, echo = FALSE}
ppmr3 <- stomach_all %>%
  mutate(numbers = log(wpredator/(wprey * sum(Nprey))),
         biomass = log(wpredator*sum(Nprey)/wprey)) %>%
  filter(Species == "Limanda limanda" )

my_mix1 <- normalmixEM(ppmr3$numbers, k = 2)
my_mix2 <- normalmixEM(ppmr3$biomass, k = 2)

#plot of numbers 
ggplot(ppmr3, aes(x = numbers)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Limanda limanda") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix1[["mu"]], #mean
    sd = my_mix1[["sigma"]], #standard deviation
    lambda = my_mix1[["lambda"]], #amplitude
    n = length(ppmr3$numbers), #sample size
    binwidth = binsize #binwidth used for histogram
  )
#plot of biomass
ggplot(ppmr3, aes(x = biomass)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Limanda limanda") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix2[["mu"]], #mean
    sd = my_mix2[["sigma"]], #standard deviation
    lambda = my_mix2[["lambda"]], #amplitude
    n = length(ppmr3$biomass), #sample size
    binwidth = binsize #binwidth used for histogram
  )
```
 
Pleuronectes platessa
```{r, echo = FALSE}
ppmr4 <- stomach_all %>%
  mutate(numbers = log(wpredator/(wprey * sum(Nprey))),
         biomass = log(wpredator*sum(Nprey)/wprey)) %>%
  filter(Species == "Pleuronectes platessa")


my_mix1 <- normalmixEM(ppmr4$numbers, k = 2)
my_mix2 <- normalmixEM(ppmr4$biomass, k = 2)

#plot of numbers 
ggplot(ppmr4, aes(x = numbers)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Pleuronectes platessa") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix1[["mu"]], #mean
    sd = my_mix1[["sigma"]], #standard deviation
    lambda = my_mix1[["lambda"]], #amplitude
    n = length(ppmr4$numbers), #sample size
    binwidth = binsize #binwidth used for histogram
  )
#plot of biomass
ggplot(ppmr4, aes(x = biomass)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Pleuronectes platessa") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix2[["mu"]], #mean
    sd = my_mix2[["sigma"]], #standard deviation
    lambda = my_mix2[["lambda"]], #amplitude
    n = length(ppmr4$biomass), #sample size
    binwidth = binsize #binwidth used for histogram
  )
```

Gadus morhua
```{r, echo = FALSE}
ppmr5 <- stomach_all %>%
  mutate(numbers = log(wpredator/(wprey * sum(Nprey))),
         biomass = log(wpredator*sum(Nprey)/wprey)) %>%
  filter(Species == "Gadus morhua" )

my_mix1 <- normalmixEM(ppmr5$numbers, k = 2)
my_mix2 <- normalmixEM(ppmr5$biomass, k = 2)

#plot of numbers 
ggplot(ppmr5, aes(x = numbers)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Gadus morhua") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix1[["mu"]], #mean
    sd = my_mix1[["sigma"]], #standard deviation
    lambda = my_mix1[["lambda"]], #amplitude
    n = length(ppmr5$numbers), #sample size
    binwidth = binsize #binwidth used for histogram
  )
#plot of biomass
ggplot(ppmr5, aes(x = biomass)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Gadus morhua") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix2[["mu"]], #mean
    sd = my_mix2[["sigma"]], #standard deviation
    lambda = my_mix2[["lambda"]], #amplitude
    n = length(ppmr5$biomass), #sample size
    binwidth = binsize #binwidth used for histogram
  )

```

Merluccius merluccius
```{r, echo = FALSE}
ppmr6 <- stomach_all %>%
  mutate(numbers = log(wpredator/(wprey * sum(Nprey))),
         biomass = log(wpredator*sum(Nprey)/wprey)) %>%
  filter(Species == "Merluccius merluccius")

my_mix1 <- normalmixEM(ppmr6$numbers, k = 2)
my_mix2 <- normalmixEM(ppmr6$biomass, k = 2)

#plot of numbers 
ggplot(ppmr6, aes(x = numbers)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Merluccius merluccius") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix1[["mu"]], #mean
    sd = my_mix1[["sigma"]], #standard deviation
    lambda = my_mix1[["lambda"]], #amplitude
    n = length(ppmr6$numbers), #sample size
    binwidth = binsize #binwidth used for histogram
  )
#plot of biomass
ggplot(ppmr6, aes(x = biomass)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Merluccius merluccius") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix2[["mu"]], #mean
    sd = my_mix2[["sigma"]], #standard deviation
    lambda = my_mix2[["lambda"]], #amplitude
    n = length(ppmr6$biomass), #sample size
    binwidth = binsize #binwidth used for histogram
  )
```

Merlangius merlangus
```{r, echo = FALSE}
ppmr7 <- stomach_all %>%
  mutate(numbers = log(wpredator/(wprey * sum(Nprey))),
         biomass = log(wpredator*sum(Nprey)/wprey)) %>%
  filter(Species == "Merlangius merlangus")

my_mix1 <- normalmixEM(ppmr7$numbers, k = 2)
my_mix2 <- normalmixEM(ppmr7$biomass, k = 2)

#plot of numbers 
ggplot(ppmr7, aes(x = numbers)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Merlangius merlangus") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix1[["mu"]], #mean
    sd = my_mix1[["sigma"]], #standard deviation
    lambda = my_mix1[["lambda"]], #amplitude
    n = length(ppmr7$numbers), #sample size
    binwidth = binsize #binwidth used for histogram
  )
#plot of biomass
ggplot(ppmr7, aes(x = biomass)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Merlangius merlangus") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix2[["mu"]], #mean
    sd = my_mix2[["sigma"]], #standard deviation
    lambda = my_mix2[["lambda"]], #amplitude
    n = length(ppmr7$biomass), #sample size
    binwidth = binsize #binwidth used for histogram
  )
```

Melanogrammus aeglefinus
```{r, echo = FALSE}
ppmr8 <- stomach_all %>%
  mutate(numbers = log(wpredator/(wprey * sum(Nprey))),
         biomass = log(wpredator*sum(Nprey)/wprey)) %>%
  filter(Species == "Melanogrammus aeglefinus")

my_mix1 <- normalmixEM(ppmr8$numbers, k = 2)
my_mix2 <- normalmixEM(ppmr8$biomass, k = 2)

#plot of numbers 
ggplot(ppmr8, aes(x = numbers)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Melanogrammus aeglefinus") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix1[["mu"]], #mean
    sd = my_mix1[["sigma"]], #standard deviation
    lambda = my_mix1[["lambda"]], #amplitude
    n = length(ppmr8$numbers), #sample size
    binwidth = binsize #binwidth used for histogram
  )
#plot of biomass
ggplot(ppmr8, aes(x = biomass)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Melanogrammus aeglefinus") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix2[["mu"]], #mean
    sd = my_mix2[["sigma"]], #standard deviation
    lambda = my_mix2[["lambda"]], #amplitude
    n = length(ppmr8$biomass), #sample size
    binwidth = binsize #binwidth used for histogram
  )
```

Eutrigla gurnardus
```{r, echo = FALSE}
ppmr9 <- stomach_all %>%
  mutate(numbers = log(wpredator/(wprey * sum(Nprey))),
         biomass = log(wpredator*sum(Nprey)/wprey)) %>%
  filter(Species == "Eutrigla gurnardus" )

my_mix1 <- normalmixEM(ppmr9$numbers, k = 2)
my_mix2 <- normalmixEM(ppmr9$biomass, k = 2)

#plot of numbers 
ggplot(ppmr9, aes(x = numbers)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Eutrigla gurnardus") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix1[["mu"]], #mean
    sd = my_mix1[["sigma"]], #standard deviation
    lambda = my_mix1[["lambda"]], #amplitude
    n = length(ppmr9$numbers), #sample size
    binwidth = binsize #binwidth used for histogram
  )
#plot of biomass
ggplot(ppmr9, aes(x = biomass)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Eutrigla gurnardus") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix2[["mu"]], #mean
    sd = my_mix2[["sigma"]], #standard deviation
    lambda = my_mix2[["lambda"]], #amplitude
    n = length(ppmr9$biomass), #sample size
    binwidth = binsize #binwidth used for histogram
  )
```

Trachurus trachurus
```{r, echo = FALSE}
ppmr10 <- stomach_all %>%
  mutate(numbers = log(wpredator/(wprey * sum(Nprey))),
         biomass = log(wpredator*sum(Nprey)/wprey)) %>%
  filter(Species == "Trachurus trachurus")

my_mix1 <- normalmixEM(ppmr10$numbers, k = 2)
my_mix2 <- normalmixEM(ppmr10$biomass, k = 2)

#plot of numbers 
ggplot(ppmr10, aes(x = numbers)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Trachurus trachurus") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix1[["mu"]], #mean
    sd = my_mix1[["sigma"]], #standard deviation
    lambda = my_mix1[["lambda"]], #amplitude
    n = length(ppmr10$numbers), #sample size
    binwidth = binsize #binwidth used for histogram
  )
#plot of biomass
ggplot(ppmr10, aes(x = biomass)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Trachurus trachurus") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix2[["mu"]], #mean
    sd = my_mix2[["sigma"]], #standard deviation
    lambda = my_mix2[["lambda"]], #amplitude
    n = length(ppmr10$biomass), #sample size
    binwidth = binsize #binwidth used for histogram
  )
```

Raja clavata
```{r, echo = FALSE}
ppmr11 <- stomach_all %>%
  mutate(numbers = log(wpredator/(wprey * sum(Nprey))),
         biomass = log(wpredator*sum(Nprey)/wprey)) %>%
  filter(Species == "Raja clavata" )

my_mix1 <- normalmixEM(ppmr11$numbers, k = 2)
my_mix2 <- normalmixEM(ppmr11$biomass, k = 2)

#plot of numbers 
ggplot(ppmr11, aes(x = numbers)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Raja clavata") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix1[["mu"]], #mean
    sd = my_mix1[["sigma"]], #standard deviation
    lambda = my_mix1[["lambda"]], #amplitude
    n = length(ppmr11$numbers), #sample size
    binwidth = binsize #binwidth used for histogram
  )
#plot of biomass
ggplot(ppmr11, aes(x = biomass)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Raja clavata") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix2[["mu"]], #mean
    sd = my_mix2[["sigma"]], #standard deviation
    lambda = my_mix2[["lambda"]], #amplitude
    n = length(ppmr11$biomass), #sample size
    binwidth = binsize #binwidth used for histogram
  )

```

Scophthalmus maximus
```{r, echo = FALSE}
ppmr12 <- stomach_all %>%
  mutate(numbers = log(wpredator/(wprey * sum(Nprey))),
         biomass = log(wpredator*sum(Nprey)/wprey)) %>%
  filter(Species == "Scophthalmus maximus")

my_mix1 <- normalmixEM(ppmr12$numbers, k = 2)
my_mix2 <- normalmixEM(ppmr12$biomass, k = 2)

#plot of numbers 
ggplot(ppmr12, aes(x = numbers)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Scophthalmus maximus") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix1[["mu"]], #mean
    sd = my_mix1[["sigma"]], #standard deviation
    lambda = my_mix1[["lambda"]], #amplitude
    n = length(ppmr12$numbers), #sample size
    binwidth = binsize #binwidth used for histogram
  )
#plot of biomass
ggplot(ppmr12, aes(x = biomass)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Scophthalmus maximus") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix2[["mu"]], #mean
    sd = my_mix2[["sigma"]], #standard deviation
    lambda = my_mix2[["lambda"]], #amplitude
    n = length(ppmr12$biomass), #sample size
    binwidth = binsize #binwidth used for histogram
  )
```

Amblyraja radiata
```{r, echo = FALSE}
ppmr13 <- stomach_all %>%
  mutate(numbers = log(wpredator/(wprey * sum(Nprey))),
         biomass = log(wpredator*sum(Nprey)/wprey)) %>%
  filter(Species == "Amblyraja radiata" )

my_mix1 <- normalmixEM(ppmr13$numbers, k = 2)
my_mix2 <- normalmixEM(ppmr13$biomass, k = 2)

#plot of numbers 
ggplot(ppmr13, aes(x = numbers)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Amblyraja radiata") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix1[["mu"]], #mean
    sd = my_mix1[["sigma"]], #standard deviation
    lambda = my_mix1[["lambda"]], #amplitude
    n = length(ppmr13$numbers), #sample size
    binwidth = binsize #binwidth used for histogram
  )
#plot of biomass
ggplot(ppmr13, aes(x = biomass)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Amblyraja radiata") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix2[["mu"]], #mean
    sd = my_mix2[["sigma"]], #standard deviation
    lambda = my_mix2[["lambda"]], #amplitude
    n = length(ppmr13$biomass), #sample size
    binwidth = binsize #binwidth used for histogram
  )
```

Pollachius virens
```{r, echo = FALSE}
ppmr14 <- stomach_all %>%
  mutate(numbers = log(wpredator/(wprey * sum(Nprey))),
         biomass = log(wpredator*sum(Nprey)/wprey)) %>%
  filter(Species == "Pollachius virens")

my_mix1 <- normalmixEM(ppmr14$numbers, k = 2)
my_mix2 <- normalmixEM(ppmr14$biomass, k = 2)

#plot of numbers 
ggplot(ppmr14, aes(x = numbers)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Pollachius virens") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix1[["mu"]], #mean
    sd = my_mix1[["sigma"]], #standard deviation
    lambda = my_mix1[["lambda"]], #amplitude
    n = length(ppmr14$numbers), #sample size
    binwidth = binsize #binwidth used for histogram
  )
#plot of biomass
ggplot(ppmr14, aes(x = biomass)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Pollachius virens") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix2[["mu"]], #mean
    sd = my_mix2[["sigma"]], #standard deviation
    lambda = my_mix2[["lambda"]], #amplitude
    n = length(ppmr14$biomass), #sample size
    binwidth = binsize #binwidth used for histogram
  )
```

Lepidorhombus whiffiagonis
```{r, echo = FALSE}
ppmr15 <- stomach_all %>%
  mutate(numbers = log(wpredator/(wprey * sum(Nprey))),
         biomass = log(wpredator*sum(Nprey)/wprey)) %>%
  filter(Species == "Lepidorhombus whiffiagonis")

my_mix1 <- normalmixEM(ppmr15$numbers, k = 2)
my_mix2 <- normalmixEM(ppmr15$biomass, k = 2)

#plot of numbers 
ggplot(ppmr15, aes(x = numbers)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Lepidorhombus whiffiagonis") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix1[["mu"]], #mean
    sd = my_mix1[["sigma"]], #standard deviation
    lambda = my_mix1[["lambda"]], #amplitude
    n = length(ppmr15$numbers), #sample size
    binwidth = binsize #binwidth used for histogram
  )
#plot of biomass
ggplot(ppmr15, aes(x = biomass)) +
  geom_histogram(binwidth = binsize) +
  ggtitle("Lepidorhombus whiffiagonis") +
  mapply(
    function(mean, sd, lambda, n, binwidth) {
      stat_function(
        fun = function(x) {
          (dnorm(x, mean = mean, sd = sd)) * n * binwidth * lambda
        }
      )
    },
    mean = my_mix2[["mu"]], #mean
    sd = my_mix2[["sigma"]], #standard deviation
    lambda = my_mix2[["lambda"]], #amplitude
    n = length(ppmr15$biomass), #sample size
    binwidth = binsize #binwidth used for histogram
  )
```

