---
title: "Reproducing Graphs"
author: "Georgina Shaw"
date: "30/11/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("data/stomach_dataset.Rdata")
library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
```

# Cod 

Only keeping the information we need 
```{r}
df <- stom_df %>%
  transmute(Species = pred_species,
            wprey = prey_weight_g,
            wpredator = pred_weight_g,
            Nprey = prey_count / n_stomachs,
            l = log(wpredator / wprey))
```
Choosing the type of fish to use, this time looking at cod 
```{r}
stomach <- df %>%
  filter(Species == "Gadus morhua",
         wprey > 0)

stomach %>% 
  group_by(Species) %>% 
  summarise(wprey_min = min(wprey),
            wprey_max = max(wprey),
            lmin = min(l),
            lmax = max(l))

stomach %>% 
  group_by(Species) %>% 
  filter(wprey == max(wprey))
```
Creating bins for the data 
```{r}
no_bins <- 30  # Number of bins
binsize <- (max(stomach$l) - min(stomach$l)) / (no_bins - 1)
breaks <- seq(min(stomach$l) - binsize/2,
              by = binsize, length.out = no_bins + 1)
```
Splitting the data into the bins that have been made
```{r}
binned_stomach <- stomach %>% 
  # bin data
  mutate(cut = cut(l, breaks = breaks, right = FALSE,
                   labels = FALSE)) %>% 
  group_by(Species, cut) %>% 
  summarise(Numbers = sum(Nprey), 
            Biomass = sum(Nprey * wprey)) %>% 
  # normalise
  mutate(Numbers = Numbers / sum(Numbers) / binsize,
         Biomass = Biomass / sum(Biomass) / binsize)  %>%
  # column for predator/prey size ratio
  mutate(l = map_dbl(cut, function(idx) breaks[idx] + binsize/2))
binned_stomach
```
We convert this into the long table format preferred by ggplot2.
```{r}
binned_stomach <- binned_stomach %>%
  gather(key = "Type", value = "Density", Numbers, Biomass)
```
                
## Histograms 
Plot the histogram that represents estimates of the normalised number density and the normalised biomass density 
```{r}
binned_stomach %>% 
  ggplot(aes(l, Density, fill = Type)) +
  geom_col(position = "dodge") +
  facet_grid(rows = vars(Species), scales = "free_y") +
  xlab("Log of Predator/Prey mass ratio") +
  expand_limits(x = c(0, 30))
```
          
## Kernal Density Estimation 
```{r}
adjust <- 1/2  # decrease bandwidth for kernel estimate
stomach <- stomach %>% 
  group_by(Species) %>% 
  mutate(weight_numbers = Nprey / sum(Nprey),
         weight_biomass = Nprey * wprey / sum(Nprey * wprey))
ggplot(stomach) +
  geom_density(aes(l, weight = weight_numbers,
                   fill = "Numbers"),
               adjust = adjust) +
  geom_density(aes(l, weight = weight_biomass,
                   fill = "Biomass"),
               adjust = adjust) +
  facet_grid(rows = vars(Species), scales = "free_y") +
  xlab("Log of predator/prey mass ratio") +
  expand_limits(x = c(0, 29))
```
         
## Gaussian Distribution fit
Plotting the normal distribution for numbers 
```{r}
weighted.sd <- function(x, w) {
  sqrt(sum(w * (x - weighted.mean(x, w))^2))
}
fit <- stomach %>% 
  summarise(mean = weighted.mean(l, weight_numbers),
            sd = weighted.sd(l, weight_numbers))
stomach %>% 
  ggplot() +
  geom_density(aes(l, weight = weight_numbers),
               fill = "#00BFC4", adjust = adjust) +
  xlab("Log of Predator/Prey mass ratio") +
  stat_function(fun = dnorm, 
                args = list(mean = fit$mean, 
                            sd = fit$sd), 
                colour = "blue") +
  expand_limits(x = c(6, 30))
```
         
Now adding biomass
```{r}
stomach %>% 
  ggplot() +
  geom_density(aes(l, weight = weight_numbers,
                   fill = "Numbers"),
               adjust = adjust) +
  geom_density(aes(l, weight = weight_biomass,
                   fill = "Biomass"),
               adjust = adjust) +
  xlab("Log of Predator/Prey mass ratio") +
  stat_function(fun = dnorm, 
                args = list(mean = fit$mean, 
                            sd = fit$sd), 
                colour = "blue") +
  stat_function(fun = dnorm, 
                args = list(mean = fit$mean - fit$sd^2, 
                            sd = fit$sd), 
                colour = "red") +
  expand_limits(x = c(0, 30))
```

```{r}
weighted.sd <- function(x, w) {
  sqrt(sum(w * (x - weighted.mean(x, w))^2))
}
fit <- stomach %>% 
  summarise(mean = weighted.mean(l, weight_numbers),
            sd = weighted.sd(l, weight_numbers))
stomach %>% 
  ggplot() +
  geom_density(aes(l, weight = weight_numbers,
                   fill = "Numbers"),
               adjust = adjust) +
  geom_density(aes(l, weight = weight_biomass,
                   fill = "Biomass"),
               adjust = adjust) +
  xlab("Log of Predator/Prey mass ratio") +
  stat_function(fun = dnorm, 
                args = list(mean = fit$mean, 
                            sd = fit$sd), 
                colour = "blue") +
  stat_function(fun = dnorm, 
                args = list(mean = fit$mean - fit$sd^2, 
                            sd = fit$sd), 
                colour = "red") +
  expand_limits(x = c(0, 30))
```
        
## Fitting the Exponential DIstribution   
Replot the histograms but with the logarithmic y-axis 
```{r}
binned_stomach %>% 
  ggplot(aes(l, Density)) +
  geom_col() +
  facet_grid(Species ~ Type, scales = "free_y") +
  xlab("Log of Predator/Prey mass ratio") +
  scale_y_continuous(trans = "log")
```
Fitting the truncated exponential distribution 
```{r}
binned_stomach %>% 
  ggplot(aes(l, Density)) +
  geom_col() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(Species ~ Type, scales = "free_y") +
  xlab("Log of Predator/Prey mass ratio") +
  scale_y_continuous(trans = "log")
```

```{r}
binned_stomach %>% 
  group_by(Species, Type) %>% 
  group_modify(~ broom::tidy(lm(log(Density) ~ l, data = .x))) %>% 
  filter(term == "l")
```
       
## Maximum Likelihood fit        
```{r}
est <- stomach %>% 
  group_by(Species) %>% 
  summarise(lbar = weighted.mean(l, weight_numbers),
            lmax = max(l),
            lmin = min(l),
            alpha = 1/(lmax - lbar))
est
```

```{r}
dnumbers <- function(l, alpha, lmax) {
  d <- as.numeric(l <= lmax)
  d[d > 0] <- dexp(lmax - l[d > 0], alpha)
  return(d)
}
dbiomass <- function(l, alpha, lmin) {
  d <- as.numeric(l >= lmin)
  d[d > 0] <- dexp(l[d > 0] - lmin, 1 - alpha)
  return(d)
}
selected_species <- "Gadus morhua"
binned_stomach %>% 
  filter(Species == selected_species) %>% 
  ggplot() +
    geom_col(aes(l, Density, fill = Type)) +
    stat_function(fun = dnumbers, 
                  args = list(alpha = est$alpha[est$Species == selected_species], 
                              lmax = est$lmax[est$Species == selected_species]), 
                  colour = "blue") +
    stat_function(fun = dbiomass, 
                  args = list(alpha = est$alpha[est$Species == selected_species], 
                              lmin = est$lmin[est$Species == selected_species]), 
                  colour = "red") +
    xlab("Log of Predator/Prey mass ratio")  +
    expand_limits(x = c(0, 30))
```
         
# Herrin 


Only keeping the information we need 
```{r}
df <- stom_df %>%
  transmute(Species = pred_species,
            wprey = prey_weight_g,
            wpredator = pred_weight_g,
            Nprey = prey_count / n_stomachs,
            l = log(wpredator / wprey))
```
Choosing the type of fish to use, this time looking at cod 
```{r}
stomach <- df %>%
  filter(Species == "Clupea harengus",
         wprey > 0)

stomach %>% 
  group_by(Species) %>% 
  summarise(wprey_min = min(wprey),
            wprey_max = max(wprey),
            lmin = min(l),
            lmax = max(l))

stomach %>% 
  group_by(Species) %>% 
  filter(wprey == max(wprey))
```
Creating bins for the data 
```{r}
no_bins <- 30  # Number of bins
binsize <- (max(stomach$l) - min(stomach$l)) / (no_bins - 1)
breaks <- seq(min(stomach$l) - binsize/2,
              by = binsize, length.out = no_bins + 1)
```
Splitting the data into the bins that have been made
```{r}
binned_stomach <- stomach %>% 
  # bin data
  mutate(cut = cut(l, breaks = breaks, right = FALSE,
                   labels = FALSE)) %>% 
  group_by(Species, cut) %>% 
  summarise(Numbers = sum(Nprey), 
            Biomass = sum(Nprey * wprey)) %>% 
  # normalise
  mutate(Numbers = Numbers / sum(Numbers) / binsize,
         Biomass = Biomass / sum(Biomass) / binsize)  %>%
  # column for predator/prey size ratio
  mutate(l = map_dbl(cut, function(idx) breaks[idx] + binsize/2))
binned_stomach
```
We convert this into the long table format preferred by ggplot2.
```{r}
binned_stomach <- binned_stomach %>%
  gather(key = "Type", value = "Density", Numbers, Biomass)
```
       
## Histograms 
Plot the histogram that represents estimates of the normalised number density and the normalised biomass density 
```{r}
binned_stomach %>% 
  ggplot(aes(l, Density, fill = Type)) +
  geom_col(position = "dodge") +
  facet_grid(rows = vars(Species), scales = "free_y") +
  xlab("Log of Predator/Prey mass ratio") +
  expand_limits(x = c(0, 30))
```
          
## Kernal Density Estimation 
```{r}
adjust <- 1/2  # decrease bandwidth for kernel estimate
stomach <- stomach %>% 
  group_by(Species) %>% 
  mutate(weight_numbers = Nprey / sum(Nprey),
         weight_biomass = Nprey * wprey / sum(Nprey * wprey))
ggplot(stomach) +
  geom_density(aes(l, weight = weight_numbers,
                   fill = "Numbers"),
               adjust = adjust) +
  geom_density(aes(l, weight = weight_biomass,
                   fill = "Biomass"),
               adjust = adjust) +
  facet_grid(rows = vars(Species), scales = "free_y") +
  xlab("Log of predator/prey mass ratio") +
  expand_limits(x = c(0, 29))
```
         
## Gaussian Distribution fit
Plotting the normal distribution for numbers 
```{r}
weighted.sd <- function(x, w) {
  sqrt(sum(w * (x - weighted.mean(x, w))^2))
}
fit <- stomach %>% 
  summarise(mean = weighted.mean(l, weight_numbers),
            sd = weighted.sd(l, weight_numbers))
stomach %>% 
  ggplot() +
  geom_density(aes(l, weight = weight_numbers),
               fill = "#00BFC4", adjust = adjust) +
  xlab("Log of Predator/Prey mass ratio") +
  stat_function(fun = dnorm, 
                args = list(mean = fit$mean, 
                            sd = fit$sd), 
                colour = "blue") +
  expand_limits(x = c(6, 30))
```
         
Now adding biomass
```{r}
stomach %>% 
  ggplot() +
  geom_density(aes(l, weight = weight_numbers,
                   fill = "Numbers"),
               adjust = adjust) +
  geom_density(aes(l, weight = weight_biomass,
                   fill = "Biomass"),
               adjust = adjust) +
  xlab("Log of Predator/Prey mass ratio") +
  stat_function(fun = dnorm, 
                args = list(mean = fit$mean, 
                            sd = fit$sd), 
                colour = "blue") +
  stat_function(fun = dnorm, 
                args = list(mean = fit$mean - fit$sd^2, 
                            sd = fit$sd), 
                colour = "red") +
  expand_limits(x = c(0, 30))
```

```{r}
weighted.sd <- function(x, w) {
  sqrt(sum(w * (x - weighted.mean(x, w))^2))
}
fit <- stomach %>% 
  summarise(mean = weighted.mean(l, weight_numbers),
            sd = weighted.sd(l, weight_numbers))
stomach %>% 
  ggplot() +
  geom_density(aes(l, weight = weight_numbers,
                   fill = "Numbers"),
               adjust = adjust) +
  geom_density(aes(l, weight = weight_biomass,
                   fill = "Biomass"),
               adjust = adjust) +
  xlab("Log of Predator/Prey mass ratio") +
  stat_function(fun = dnorm, 
                args = list(mean = fit$mean, 
                            sd = fit$sd), 
                colour = "blue") +
  stat_function(fun = dnorm, 
                args = list(mean = fit$mean - fit$sd^2, 
                            sd = fit$sd), 
                colour = "red") +
  expand_limits(x = c(0, 30))
```
        
## Fitting the Exponential DIstribution   
Replot the histograms but with the logarithmic y-axis 
```{r}
binned_stomach %>% 
  ggplot(aes(l, Density)) +
  geom_col() +
  facet_grid(Species ~ Type, scales = "free_y") +
  xlab("Log of Predator/Prey mass ratio") +
  scale_y_continuous(trans = "log")
```
Fitting the truncated exponential distribution 
```{r}
binned_stomach %>% 
  ggplot(aes(l, Density)) +
  geom_col() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(Species ~ Type, scales = "free_y") +
  xlab("Log of Predator/Prey mass ratio") +
  scale_y_continuous(trans = "log")
```

```{r}
binned_stomach %>% 
  group_by(Species, Type) %>% 
  group_modify(~ broom::tidy(lm(log(Density) ~ l, data = .x))) %>% 
  filter(term == "l")
```
       
## Maximum Likelihood fit        
```{r}
est <- stomach %>% 
  group_by(Species) %>% 
  summarise(lbar = weighted.mean(l, weight_numbers),
            lmax = max(l),
            lmin = min(l),
            alpha = 1/(lmax - lbar))
est
```

```{r}
dnumbers <- function(l, alpha, lmax) {
  d <- as.numeric(l <= lmax)
  d[d > 0] <- dexp(lmax - l[d > 0], alpha)
  return(d)
}
dbiomass <- function(l, alpha, lmin) {
  d <- as.numeric(l >= lmin)
  d[d > 0] <- dexp(l[d > 0] - lmin, 1 - alpha)
  return(d)
}
selected_species <- "Clupea harengus"
binned_stomach %>% 
  filter(Species == selected_species) %>% 
  ggplot() +
    geom_col(aes(l, Density, fill = Type)) +
    stat_function(fun = dnumbers, 
                  args = list(alpha = est$alpha[est$Species == selected_species], 
                              lmax = est$lmax[est$Species == selected_species]), 
                  colour = "blue") +
    stat_function(fun = dbiomass, 
                  args = list(alpha = est$alpha[est$Species == selected_species], 
                              lmin = est$lmin[est$Species == selected_species]), 
                  colour = "red") +
    xlab("Log of Predator/Prey mass ratio")  +
    expand_limits(x = c(0, 30))
```